---
output: rmarkdown::github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figures/DR_Kessenbrock-", 
  echo = TRUE)
``` 
# Analysis of integrated breast epithelial atlas
```{r}
library(Seurat)
library(ggplot2)
```

Nguyen, et. al used 10X Chromium single-cell RNA sequencing to profile mammary epithelial cells of 4 individuals. The raw count data from this experiment were [quality filtered with custom R scripts](https://github.com/jasenjackson/singlecell/blob/master/QC_Kessenbrock.md), normalized for sequencing depth using scTransform and [integrated](https://github.com/jasenjackson/singlecell/blob/master/integrate.R) using Seurat's reference-based integration method. 
```{r, fig.width=10, fig.height=5, warning=FALSE}
BreastAtlas <- readRDS("/projects/b1101/Jasen/data/BreastAtlas.integrated.rds")

# Perform louvain algorithm via the Seurat package to identify cell clusters
BreastAtlas <- FindNeighbors(BreastAtlas, dims = c(1:30), verbose = FALSE)
BreastAtlas <- FindClusters(BreastAtlas, verbose = FALSE)

# Plot by Individual and by Seurat cluster
p1 <- DimPlot(BreastAtlas, reduction = "umap", group.by = "individual", pt.size=0.1) + DarkTheme()
p2 <- DimPlot(BreastAtlas, reduction = "umap", group.by = "seurat_clusters", pt.size=0.1) + DarkTheme()
gridExtra::grid.arrange(p1,p2, ncol=2)
```

Nguyen, et. al identified three primary types of mammary cells: Basal, Luminal 1 (L1) and Luminal 2 (L2). KRT18 was identified as a luminal marker, whereas KRT14 was identified as a basal marker. SLPI and ANKRD30A were identified as marker genes for Luminal 1 and Luminal2 respectively. CD74, AGR2, LTF, SAA2, KRT23, ACTA2, KRT17, and IL24 were identified as within-type marker genes as well. Generally, these marker gene expression patterns were reproducible following integration across datasets. 
```{r, fig.width = 12, fig.height = 6}
featureMap.df <- FetchData(object = BreastAtlas, vars = c("UMAP_1", "UMAP_2",
                                                          "KRT18", "KRT14", "SLPI", "ANKRD30A",
                                                          "CD74", "AGR2", "LTF", "SAA2",
                                                          "KRT23","ACTA2", "KRT17", "IL24"))

p1 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = KRT18)) + geom_point(size=0.1) + scale_color_viridis_c()
p2 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = KRT14)) + geom_point(size=0.1) + scale_color_viridis_c()
p3 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = SLPI)) + geom_point(size=0.1) + scale_color_viridis_c()
p4 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = ANKRD30A)) + geom_point(size=0.1) + scale_color_viridis_c()
p5 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = CD74)) + geom_point(size=0.1) + scale_color_viridis_c()
p6 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = AGR2)) + geom_point(size=0.1) + scale_color_viridis_c()
p7 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = LTF)) + geom_point(size=0.1) + scale_color_viridis_c()
p8 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = SAA2)) + geom_point(size=0.1) + scale_color_viridis_c()
p9 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = KRT23)) + geom_point(size=0.1) + scale_color_viridis_c()
p10 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = ACTA2)) + geom_point(size=0.1) + scale_color_viridis_c()
p11 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = KRT17)) + geom_point(size=0.1) + scale_color_viridis_c()
p12 <- ggplot(featureMap.df, aes(UMAP_1, UMAP_2, color = IL24)) + geom_point(size=0.1) + scale_color_viridis_c()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 4)
```

```{r}
# stash identity
BreastAtlas[["sample_origin"]] <- BreastAtlas[["orig.ident"]]
Idents(object = BreastAtlas) <- BreastAtlas[["seurat_clusters"]]

Misc(object = BreastAtlas, slot = "cluster0.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "0", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster1.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "1", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster2.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "2", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster3.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "3", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster4.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "4", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster5.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "5", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster6.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "6", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster7.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "7", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster8.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "8", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster9.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "9", ident.2 = NULL, only.pos = TRUE), "gene")
Misc(object = BreastAtlas, slot = "cluster10.markers") <- tibble::rownames_to_column(FindMarkers(BreastAtlas, ident.1 = "10", ident.2 = NULL, only.pos = TRUE), "gene")

#Cluster 0
paste(BreastAtlas@misc$cluster0.markers$gene, collapse=",")

#Cluster 1
paste(BreastAtlas@misc$cluster1.markers$gene, collapse=",")

#Cluster 2
paste(BreastAtlas@misc$cluster2.markers$gene, collapse=",")

#Cluster 3
paste(BreastAtlas@misc$cluster3.markers$gene, collapse=",")

#Cluster 4
paste(BreastAtlas@misc$cluster4.markers$gene, collapse=",")

#Cluster 5
paste(BreastAtlas@misc$cluster5.markers$gene, collapse=",")

#Cluster 6
paste(BreastAtlas@misc$cluster6.markers$gene, collapse=",")

#Cluster 7
paste(BreastAtlas@misc$cluster7.markers$gene, collapse=",")

#Cluster 8
paste(BreastAtlas@misc$cluster8.markers$gene, collapse=",")

#Cluster 9
paste(BreastAtlas@misc$cluster9.markers$gene, collapse=",")

#Cluster 10
paste(BreastAtlas@misc$cluster10.markers$gene, collapse=",")
```

```{r}
saveRDS(BreastAtlas, file="/projects/b1101/Jasen/data/BreastAtlas.rds")
```
  