---
title: "QC Filtering of Breast Epithelial Atlas"
author: "Jasen Jackson"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
# import packages and functions
setwd('/projects/b1101/Jasen')
source('singlecell/QCfunctions.R')

library(matrixStats)
library(scales)
library(ggplot2)
library(dplyr)
library(stringr)
```

Import the count matrix for individual 4 from the Breast Epithelial atlas (Nguyen, et. al 2018; doi: 10.1038/s41467-018-04334-1). The rows are genes and the columns are cells.

```{r, cache=TRUE}
Ind4.df <- read.table(file = "/projects/b1101/Jasen/data/Breast_scRNA_Kessenbrock_Ind4.txt", header = TRUE, row.names=1)
head(Ind4.df[,1:4])
```

## Calculate statistics for each gene and cell. 
The raw count, mean and standard deviation are computed by the stats_by_gene() function. The gene count, read (umi) count and percent of reads mapped to mitochondrial genes ('percent.mt') are calculated by default using the stats_by_cell() function. If the matrix computed by the stats_by_gene() function is included then 'percent.cg', an additional QC metric, is calculated.
```{r}
Ind4.genestats <- stats_by_gene(Ind4.df)
Ind4.genestats[1:4,]
```

The stats_by_gene() function calculates the % of cells expressing each gene, and stats_by_cell() uses this metric to define a set of housekeeping genes or common genes (expressed in >95% of cells). The % of housekeeping (common) genes expressed in each cell is then saved in the 'percent.cg' column. 
```{r}
Ind4.cellstats <- stats_by_cell(Ind4.df, Ind4.genestats, common_genes=TRUE)
```
```{r}
Ind4.cellstats[1:4,]
```