---
output: rmarkdown::github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figures/QC_filtering-", 
  echo = TRUE)
```

# QC Filtering of Breast Epithelial Atlas
```{r, message=FALSE}
# import packages and functions
setwd('/projects/b1101/Jasen')
source('singlecell/QCfunctions.R')

library(matrixStats)
library(scales)
library(ggplot2)
library(stringr)
library(dplyr)
```

## Import the count matrix
Import the count matrix forrreal individual 4 from the Breast Epithelial atlas (Nguyen, et. al 2018; doi:[10.1038/s41467-018-04334-1](https://www.nature.com/articles/s41467-018-04334-1). <br /> <br /> The rows are genes and the columns are cells.

```{r}
Ind4.df <- read.table(file = "/projects/b1101/Jasen/data/Breast_scRNA_Kessenbrock_Ind4.txt", header = TRUE, row.names=1)
head(Ind4.df[,1:4])
```

## Calculate statistics for each gene and cell. 
The mean and standard deviation and % of cells expressing each gene are computed by the stats_by_gene() function. 
```{r}
Ind4.genestats <- stats_by_gene(Ind4.df)
Ind4.genestats[1:4,]
```

The gene count, read (umi) count and percent of reads mapped to mitochondrial genes ('percent.mt') are calculated by default using the stats_by_cell() function. If the matrix computed by the stats_by_gene() function is included then 'percent.cg', an additional QC metric, is calculated.

The stats_by_gene() function calculates the % of cells expressing each gene, and stats_by_cell() uses this metric to define a set of housekeeping genes or common genes (expressed in >95% of cells). The % of housekeeping (common) genes expressed in each cell is then saved in the 'percent.cg' column. 
```{r}
Ind4.cellstats <- stats_by_cell(Ind4.df, Ind4.genestats, common_genes=TRUE)
```
```{r}
Ind4.cellstats[1:4,]
```

Typically, cells are filtered jointly by gene count, umi count and percentage of mitochondrial genes. Low quality cells tend to have low values for each of these features. 
```{r}
ggplot(Ind4.cellstats, aes(umi_count, gene_count, color = percent.mt)) +
       geom_point(size=0.7) +
       scale_color_viridis_c()
```

Percentage of housekeeping genes (common genes percentage) may be a more precise alternative to filtering on mitochondrial percentage. We can plot % of common genes detected in each cell to determine a cutoff. 
```{r}
ggplot(Ind4.cellstats, aes(reorder(cell, percent.cg), percent.cg))  + 
       geom_point(size=0.7) 
```
It looks like there is an elbow around 85%. We can compare this cutoff to gene count and umi_count. 
```{r}
ggplot(Ind4.cellstats, aes(umi_count, gene_count, color=percent.cg)) +
  geom_point(size=0.7) +
  scale_color_viridis_c(direction = -1)
```

```{r}
```
