---
output: rmarkdown::github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figures/QC_Polyak-", 
  echo = TRUE)
```
# QC of cancer cell line data (scRNA-seq) from the [Polyak lab](https://www.polyaklab.dfci.harvard.edu/)
```{r, message=FALSE}
# import packages and functions
setwd('/projects/b1101/Jasen')
source('singlecell/QCfunctions.R')

library(matrixStats)
library(scales)
library(ggplot2)
library(stringr)
library(dplyr)
library(MASS)
library(Seurat)
```

## MCF7 #1 ([GSM2813220](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM2813220)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
MCF7_1.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM2813220_MCF7.tsv", header = TRUE, row.names=1)
ncol(MCF7_1.mtx); nrow(MCF7_1.mtx)

# Calculate and plot QC statistics
MCF7_1.cellstats <- QCstats(MCF7_1.mtx, verbose=FALSE)
QCplot(MCF7_1.cellstats, hline=1000, vline=10000) 
```
\n
1175 cells, 7,738 genes.
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
MCF7_1.meta <- QCseurat(MCF7_1.cellstats) # formats cellstats matrix for Seurat
MCF7_1 <- CreateSeuratObject(count=MCF7_1.mtx, project="MCF7_1", meta.data=MCF7_1.meta)
MCF7_1 <- subset(x = MCF7_1, subset = percent.cg > 90 & gene_count > 1000 & umi_count < 10000 & percent.mt < 50)
ncol(MCF7_1.mtx) - ncol(MCF7_1); ((ncol(MCF7_1.mtx)-ncol(MCF7_1)) / ncol(MCF7_1.mtx)) * 100
```
\n
62 cells (5.276%) removed. 
```{r}
# remove raw data from memory
rm(MCF7_1.cellstats); rm(MCF7_1.mtx); rm(MCF7_1.meta)
```

## MCF7 #2 ([GSM2813221](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM2813221)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
MCF7_2.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM2813221_pMCF7.tsv", header = TRUE, row.names=1)
ncol(MCF7_2.mtx); nrow(MCF7_2.mtx)

# Calculate and plot QC statistics
MCF7_2.cellstats <- QCstats(MCF7_2.mtx, verbose=FALSE)
QCplot(MCF7_2.cellstats, vline = 7000) 
```
\n
927 cells, 7,738 genes detected. Low over-all gene count per cell, so this will not be used to filter these cells. 
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
MCF7_2.meta <- QCseurat(MCF7_2.cellstats) # formats cellstats matrix for Seurat
MCF7_2 <- CreateSeuratObject(count=MCF7_2.mtx, project="MCF7_2", meta.data=MCF7_2.meta)
MCF7_2 <- subset(x = MCF7_2, subset = percent.cg > 90 & percent.mt < 50 & umi_count < 7000)
ncol(MCF7_2.mtx) - ncol(MCF7_2); ((ncol(MCF7_2.mtx)-ncol(MCF7_2)) / ncol(MCF7_2.mtx)) * 100
```
\n
48 cells (~5.2%) were removed. 
```{r}
# remove raw data from memory
rm(MCF7_2.cellstats); rm(MCF7_2.mtx); rm(MCF7_2.meta)
```

## SUM159 ([GSM2813221](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM2813221)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
SUM159.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM2813221_SUM159.tsv", header = TRUE, row.names=1)
ncol(SUM159.mtx); nrow(SUM159.mtx)

# Calculate and plot QC statistics
SUM159.cellstats <- QCstats(SUM159.mtx, verbose=FALSE)
QCplot(SUM159.cellstats, hline=1000, vline=9000) 
```
\n
2,276 cells, 7,738 genes were detected. May be able to filter up to >95% percent.cg expressed! 
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
SUM159.meta <- QCseurat(SUM159.cellstats) # formats cellstats matrix for Seurat
SUM159 <- CreateSeuratObject(count=SUM159.mtx, project="SUM159", meta.data=SUM159.meta)
SUM159 <- subset(x = SUM159, subset = percent.cg > 90 & gene_count > 1000 & umi_count < 9000 & percent.mt < 50)
ncol(SUM159.mtx) - ncol(SUM159); ((ncol(SUM159.mtx)-ncol(SUM159)) / ncol(SUM159.mtx)) * 100
```
\n
43 cells (1.9%) were removed. 
```{r}
# remove raw data from memory
rm(SUM159.cellstats); rm(SUM159.mtx); rm(SUM159.meta)
```

## BT474 ([GSM3329993](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM3329993)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
BT474.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM3329993_BT474.tsv", header = TRUE, row.names=1)
ncol(BT474.mtx); nrow(BT474.mtx)
head(BT474.mtx[,1:3])
# Calculate and plot QC statistics
BT474.cellstats <- QCstats(BT474.mtx, verbose=FALSE)
QCplot(BT474.cellstats,hline=950, vline=9000) 
```
\n
3,562 cells, 8,019 genes detected. Very high overall mt %, not filtered out. 
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
BT474.meta <- QCseurat(BT474.cellstats) # formats cellstats matrix for Seurat
BT474 <- CreateSeuratObject(count=BT474.mtx, project="BT474", meta.data=BT474.meta)
BT474 <- subset(x = BT474, subset = percent.cg > 90 & gene_count > 950 & umi_count < 9000)
ncol(BT474.mtx) - ncol(BT474); ((ncol(BT474.mtx)-ncol(BT474)) / ncol(BT474.mtx)) * 100
```
\n
78 cells (~2.2%) were removed. 
```{r}
# remove raw data from memory
rm(BT474.cellstats); rm(BT474.mtx); rm(BT474.meta)
```

## T47D ([GSM3329993](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM3329993)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
T47D.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM3329993_T47D.tsv", header = TRUE, row.names=1)
ncol(T47D.mtx); nrow(T47D.mtx)
head(T47D.mtx[,1:3])

# Calculate and plot QC statistics
T47D.cellstats <- QCstats(T47D.mtx, verbose=FALSE)
QCplot(T47D.cellstats, vline=8000) 
```
\n
2,530 cells, 8,019 genes detected. High percent.mt, low gene count. 
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
T47D.meta <- QCseurat(T47D.cellstats) # formats cellstats matrix for Seurat
T47D <- CreateSeuratObject(count=T47D.mtx, project="T47D", meta.data=T47D.meta)
T47D <- subset(x = T47D, subset = percent.cg > 90 & umi_count < 8000 & percent.mt < 50)
ncol(T47D.mtx) - ncol(T47D); ((ncol(T47D.mtx)-ncol(T47D)) / ncol(T47D.mtx)) * 100
```
\n
104 cells (~4.1%) removed.

```{r}
# remove raw data from memory
rm(T47D.cellstats); rm(T47D.mtx); rm(T47D.meta)
```

## ZR75 ([GSM3329993](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM3329993)) 
```{r fig.width=14, fig.height=10, message=FALSE}
# Import the count matrix
ZR75.mtx <- read.table(file = "/projects/b1101/Jasen/data/GSM3329993_ZR75.tsv", header = TRUE, row.names=1)
ncol(ZR75.mtx); nrow(ZR75.mtx)
head(ZR75.mtx[,1:5])
# Calculate and plot QC statistics
ZR75.cellstats <- QCstats(ZR75.mtx, verbose=FALSE)
QCplot(ZR75.cellstats, hline = 900, vline = 12000) 
```
\n 
2,811 cells, 8,019 genes detected. High percent.mt
```{r, warning=FALSE}
# Load count matrix and metadata into Seurat object & filter
ZR75.meta <- QCseurat(ZR75.cellstats) # formats cellstats matrix for Seurat
ZR75 <- CreateSeuratObject(count=ZR75.mtx, project="ZR75", meta.data=ZR75.meta)
ZR75 <- subset(x = ZR75, subset = percent.cg > 90 & gene_count > 900 & umi_count < 12000 & percent.mt < 50)
ncol(ZR75.mtx) - ncol(ZR75); ((ncol(ZR75.mtx)-ncol(ZR75)) / ncol(ZR75.mtx)) * 100
```
\n 
109 cells (~3.9%) were removed. 

```{r}
# remove raw data from memory
rm(ZR75.cellstats); rm(ZR75.mtx); rm(ZR75.meta)
```

## Merging cell lines with Breast Epithelial Atlas

Each cell line was first merged into an aggregate 'BreastCancerAtlas' object. The datasets will be split by 'sample_origin' before reference-based integration. 
```{r}
BreastCancerAtlas <- merge(x = MCF7_1, y = list(MCF7_2, SUM159, T47D, ZR75, BT474))
rm(MCF7_1, MCF7_2, SUM159, T47D, ZR75, BT474)
ncol(BreastCancerAtlas); nrow(BreastCancerAtlas)
# create 'malignant' & 'is_from_cell_line' boolean column for integration
malignant <- rep(TRUE, ncol(BreastCancerAtlas))
BreastCancerAtlas[["malignant"]] <- malignant

cell_line <- rep(TRUE, ncol(BreastCancerAtlas))
BreastCancerAtlas[["cell_line"]] <- cell_line

#remove default Seurat columns to reduce file size
BreastCancerAtlas[["sample_origin"]] <- BreastCancerAtlas[["orig.ident"]]
BreastCancerAtlas[["nCount_RNA"]] <- NULL
BreastCancerAtlas[["nFeature_RNA"]] <- NULL
head(BreastCancerAtlas[[]])
```

The integrated breast epithelial dataset was then joined with the breast cancer cell lines. 
```{r}
BreastAtlas.integrated <- readRDS(file="/projects/b1101/Jasen/data/BreastAtlas.integrated.rds") 
ncol(BreastAtlas.integrated); nrow(BreastAtlas.integrated)

BreastAtlas.integrated[["sample_origin"]] <- BreastAtlas.integrated[["individual"]]
BreastAtlas.integrated[["individual"]] <- NULL

# set malignant & cell_line columns
malignant <- rep(FALSE, ncol(BreastAtlas.integrated))
BreastAtlas.integrated[["malignant"]] <- malignant

cell_line <- rep(FALSE, ncol(BreastAtlas.integrated))
BreastAtlas.integrated[["cell_line"]] <- cell_line

head(BreastAtlas.integrated[[]])

# Merge non-malignant reference cells 'under' cell line data
BreastCancerAtlas <- merge(x = BreastCancerAtlas, y = BreastAtlas.integrated)
ncol(BreastCancerAtlas); nrow(BreastCancerAtlas)
# remove redundant columns
BreastCancerAtlas[["nCount_RNA"]] <- NULL
BreastCancerAtlas[["nFeature_RNA"]] <- NULL

head(BreastCancerAtlas[[]])
tail(BreastCancerAtlas[[]])
saveRDS(BreastCancerAtlas, file="/projects/b1101/Jasen/data/BreastCancerAtlas.rds")
```



